---
bibliography: references.bib
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
format:
  html:
    code-fold: true
---

```{r load-packages}
#| message: false
#| warning: false
library(tidyverse)
library(here)
library(knitr)
library(rriskDistributions) # to fit distributions
library(RColorBrewer)
library(msm) # for truncated normal 
library(gt) # for complex tables
library(MASS) # for colinarity
library(tibble)
library(dplyr)
library(tidyr)
library(MASS)
library(ggcorrplot)
library(patchwork)
conflicted::conflicts_prefer(dplyr::filter)
```

```{r fix-seed}
set.seed(1991)
```

# Covariate simulation

## Objective

```{r}
n_patient <- 625 # number of patients simulated per cohort
```

```{r all-papers-cov-data, warning=FALSE, message=FALSE}
#| message: false
here::i_am("Simulations/covariate_simulation.qmd")

# Create folder to store published figures
if (!dir.exists(here("Figures"))) {
  dir.create(here("Figures"))
}

cov_data <- read_csv(here("Simulations/cov_distrib_from_papers.csv"))
```

From each identified paper, we want to simulate a cohort of patients. For each cohort we want to simulate `r n_patient` patients. We
define a patient by a vector of covariate values. The covariates of
interest are :

-   WT : Body weight (kg)
-   CRCL : Creatinine clearance (mL/min)
-   ICU : Is the patient an ICU patient (=1) or not (=0)
-   BURN : Is the patient a burn patient (=1) or not (=0)
-   OBESE : Is the patient a obese - BMI \> 30 kg/m\^2 (=1) or not (=0)
-   SEX : Male (=0) or female (=1)
-   HT : height (cm)
-   AGE : age (years)

To simulate covariate values we use information on the distribution of
the covariates taken from the included papers.

We included 4 papers :

-   @Carlier2013-bq

-   @fournier2018

-   @mellon2020

-   @rambaud2020

One issue with the covariate data reported in these papers is that correlations 
between covariates are not reported and simulating without correlations runs 
the risk of simulating improbable covariate values.
We thus obtained a clinical dataset composed of more than 16000 ICU patients
(@Johnson2023-zo) from which we could derive correlations between covariates of interest. Age, serum creatinine, body weight, sex, height and
BMI were obtained for these patients and the creatinine clearance was
calculated (in mL/min for Cockroft and Gault and CKD-EPI and in
mL/min/1.73 m2 for MDRD (to correspond to Mellon)). A correlation matrix
was calculated for each of the respective model development cohorts separately for females and males.


This correlation matrix is used to simulate correlations between the
simulated WT, CRCL, AGE, and HT/BMI if necessary to convert CRCL to
serum creatinine. (CRCL_MDRD_norm means CRCL in mL/min/1.73 m² used for
Mellon and CRCL_MDRD in mL/min.). Different correlation matrices are used based on sex. This way two different mutltivariate distributions are simulated based on the two matrices with a number of patients that respects the original article proportions.

We can find correlations for the log-transforms of certain variables. For the covariates where a log-normal distribution fits better the log mean and standard deviation of this distribution will be used to sample log values with a multivariate normal distribution. At the end, the log variables will be retransformed to normal ones.

## @Carlier2013-bq

### Covariate distribution

All patients are ICU patients and we will assume that they are not burn
and not obese patients. Creatinine clearance was obtained using the
measured urinary equation. Thus for all patients ICU = 1, BURN = 0.

As discussed in the article, renal replacement therapy is an exclusion
criterion, therefore the minimum value of creatinine clearance is set to
10 mL/min, any simulated value below 10 mL/min will be resampled. CRCL was calculated using the measured urinary equation.

For WT and CRCL here are the data from the paper :

```{r carlier-cov-data-paper, warning=FALSE, message=FALSE}
cov_data_carlier <- cov_data |> 
  filter(Paper == "Carlier_2013")

cov_data_carlier |> 
  filter(Covariate %in% c("WT","CRCL","AGE","BMI")) |> 
  kable()
```

```{r carlier-cor-matrix, warning=FALSE, message=FALSE}
cor_matrix_Carlier_F <- readRDS(here("Simulations/cor_matrix_Carlier_F.rds"))
print(cor_matrix_Carlier_F)
ggcorrplot(cor_matrix_Carlier_F, lab = TRUE, title = "Sex: female")

cor_matrix_Carlier_M <- readRDS(here("Simulations/cor_matrix_Carlier_M.rds"))
print(cor_matrix_Carlier_M)
ggcorrplot(cor_matrix_Carlier_M, lab = TRUE, title = "Sex: male")
```

We are going to fit normal and log-normal distributions to the observed
quantiles and select the best fitting one

```{r functions-fit-cov-distrib, warning=FALSE, message=FALSE}

compute_cv_from_sd_lnorm <- function(sdlog){
  #' Compute the geometric coefficient of variation from the standard deviation of a lognormal distribution
  #'
  #' @param sdlog numeric. standard deviation of the log transformed variable
  #' 
  #' @return numeric. geometric coefficient of variation
  sqrt(exp(sdlog^2)-1)
}

extract_cov_param <- function(cov_data, cov_name) {
  
  #' Get covariate quantiles from covariate characteristics table
  #'
  #' @param cov_data tibble containing the covariate information
  #' @param cov_name character. name of the covariate to select, comes from the Covariate column of cov_data.
  #' @return list with :
  #'  - p : numeric vector of probabilities
  #'  - q : numeric vector of quantiles
  #'  - cov_name : character string containing the covariate name
  #'  - cov_unit : character string containing the unit of the covariate
  
  
  cov_filtered_df <- cov_data |>
    filter(Covariate == cov_name) |>
    pivot_longer(
      cols = c(Median, Q1, Q3, Min, Max), # if the format of the csv columns ever changes, this will need to change
      names_to = "p",
      values_to = "q"
    ) |>
    mutate(p = case_match(p, "Median" ~ 0.5, "Q1" ~ 0.25, "Q3" ~ 0.75, "Min" ~ 0.01, "Max" ~ 0.99)) |>
    # if the format of the csv columns ever changes, the code above will need to change
    filter(!is.na(q)) |>
    group_by(Paper) |>
    arrange(p, .by_group = TRUE) |>
    ungroup()
  
  list(
    p = cov_filtered_df$p,
    q = cov_filtered_df$q,
    cov_name = unique(cov_filtered_df$Covariate),
    cov_unit = unique(cov_filtered_df$Unit)
  )
  
  
}

evaluate_distrib <- function(p, q, cov_name, cov_unit, min_plot, max_plot,tol = 0.001) {
  #' Fit normal and lognormal distribution to observed quantiles
  #' @param p numeric vector of probabilities
  #' @param q numeric vector of quantiles
  #' @param cov_name character string containing the covariate name
  #' @param cov_unit character string containing the unit of the covariate
  #' @param min_plot minimal value of the covariate for which the distribution should be plotted
  #' @param max_plot maximal value of the covariate for which the distribution should be plotted
  #' @param tol numeric. tolerance value for the fitting algorithm, default is 0.001
  #' @return list with :
  #'  - plot : ggplot2 showing the fitted distributions and the observed covariate
  #'  - lnorm_par : named numeric vector with the lognormal distribution parameters
  #'  - norm_par : named numeric vector with the normal distribution parameters
  
  
  lnorm_par <- get.lnorm.par(p = p, q = q, plot = FALSE, show.output = FALSE,tol = tol)
  norm_par <- get.norm.par(p = p, q = q, plot = FALSE, show.output = FALSE,tol=tol)
  
  #create a dataset with the true percentiles to be able to show them on the plot
  true_percentiles <- tibble(p, q) |> 
    mutate(label = case_match(p, 
                              0 ~ "True min",
                              0.25 ~ "True Q1",
                              0.5 ~ "True median",
                              0.75 ~ "True Q3",
                              1 ~ "True max")) 
  
  #create a dataset with the percentiles of the fitted distributiosn to be able to show them on the plot
  distrib_percentiles <-  tibble(
    probability = p,
    lnorm = qlnorm(probability, meanlog = lnorm_par["meanlog"], sdlog =
                     lnorm_par["sdlog"]),
    norm = qnorm(probability, mean = norm_par["mean"], sd = norm_par["sd"])
  ) |>
    pivot_longer(cols = c(lnorm, norm), names_to = "distrib") |>
    mutate(density = case_match(
      distrib,
      "norm" ~ dnorm(value, mean = norm_par["mean"], sd =
                       norm_par["sd"]),
      "lnorm" ~ dlnorm(value, meanlog = lnorm_par["meanlog"], sdlog =
                         lnorm_par["sdlog"])
    )) |> 
    mutate(label = case_match(probability, 
                              0 ~ "Min",
                              0.25 ~ "Q1",
                              0.5 ~ "Median",
                              0.75 ~ "Q3",
                              1 ~ "Max")) 
  
  
  #simulate 1000 datapoints from the fitted distributions to show on the plot
  param_data <- tibble(
    variable = seq(min_plot, max_plot, length.out = 1000),
    lnorm = dlnorm(variable, meanlog = lnorm_par["meanlog"], sdlog = lnorm_par["sdlog"]),
    norm = dnorm(variable, mean = norm_par["mean"], sd = norm_par["sd"])
  ) |>
    pivot_longer(cols = c(lnorm, norm), names_to = "distrib")
  
  
  plot <- ggplot(param_data, aes(x = variable, y = value, color = distrib)) +
    geom_vline(data = true_percentiles, mapping = aes(xintercept = q),lty = "dashed") +
    geom_line() +
    geom_segment(data = distrib_percentiles, aes(x = value, y = 0, yend =
                                                   density)) +
    theme_bw() +
    scale_color_brewer("Distribution", palette = "Dark2") +
    ylab("Density") +
    xlab(paste0(cov_name, " (", cov_unit, ")")) +
    annotate(
      geom = "text",
      label = paste0(
        "N ~ (",
        signif(norm_par["mean"], 3),
        ", ",
        signif(norm_par["sd"], 3),
        ")\n",
        "LN ~ (",
        signif(exp(lnorm_par["meanlog"]), 3),
        ", ",
        signif(compute_cv_from_sd_lnorm(lnorm_par["sdlog"]), 3) * 100,
        " %)"
      ),
      x = min(param_data$variable) + 0.99 * diff(range(param_data$variable)),
      y = min(param_data$value) + 0.99 * diff(range(param_data$value)),
      hjust = 1,
      vjust = 1
    )+
    geom_text(
      data = true_percentiles,
      mapping = aes(x = q,label = label,color=NULL),
      show.legend = FALSE,
      y = min(param_data$value) + 0.95 * diff(range(param_data$value)),
      hjust = 1,
      vjust = -1,
      angle = 90
    )+
    geom_text(
      data = distrib_percentiles,
      mapping = aes(x = value, y=density,label = label),
      show.legend = FALSE,
      hjust = 1,
      vjust = -1,
      angle = 90
    )
  
  list(plot = plot,
       lnorm_par = lnorm_par,
       norm_par = norm_par)
  
}

fit_cov_wrapper <- function(cov_data, cov_name, min_plot, max_plot, tol = 0.001) {
  
  #' Wrapper to get covariate quantiles from covariate characteristics table and fit normal and lognormal distribution
  #' to observed quantiles 
  #'
  #' @param cov_data tibble containing the covariate information
  #' @param cov_name character. name of the covariate to select, comes from the Covariate column of cov_data.
  #' @param min_plot minimal value of the covariate for which the distribution should be plotted
  #' @param max_plot maximal value of the covariate for which the distribution should be plotted
  #' @param tol numeric. tolerance value for the fitting algorithm, default is 0.001
  #' @return list with :
  #'  - plot : ggplot2 showing the fitted distributions and the observed covariate
  #'  - lnorm_par : named numeric vector with the lognormal distribution parameters
  #'  - norm_par : named numeric vector with the normal distribution parameters
  
  
  cov_df <- extract_cov_param(cov_data, cov_name)
  
  evaluate_distrib(
    p = cov_df$p,
    q = cov_df$q,
    min_plot = min_plot,
    max_plot = max_plot,
    cov_name = cov_df$cov_name,
    cov_unit = cov_df$cov_unit,
    tol = tol
  )
  
}
```

```{r fit-carlier-crcl, warning=FALSE, message=FALSE}
#| warning: false

carlier_crcl <- fit_cov_wrapper(
  cov_data = cov_data_carlier,
  cov_name = "CRCL",
  min_plot = 0,
  max_plot = 250
)

carlier_crcl$plot

```

Normal distribution seems better than log-normal.

```{r fit-carlier-wt, warning=FALSE, message=FALSE}
#| warning: false
carlier_wt <- fit_cov_wrapper(
  cov_data = cov_data_carlier,
  cov_name = "WT",
  min_plot = 40,
  max_plot = 120
)

carlier_wt$plot
```
No discernable difference in fit. We have to use the log-normal
distribution to be able to calculate height.

```{r fit-carlier-age, warning=FALSE, message=FALSE}
#| warning: false
carlier_age <- fit_cov_wrapper(
  cov_data = cov_data_carlier,
  cov_name = "AGE",
  min_plot = 0,
  max_plot = 120
)

carlier_age$plot
```

No discernible difference in fit. We will use the normal distribution as generally, age is normally distributed.

```{r fit-carlier-bmi, warning=FALSE, message=FALSE}
#| warning: false
carlier_bmi <- fit_cov_wrapper(
  cov_data = cov_data_carlier,
  cov_name = "BMI",
  min_plot = 0,
  max_plot = 40,
  tol=0.002
)

carlier_bmi$plot
```

Tolerance was increased to 0.002. No discernible difference in fit. We have to use the log-normal distribution to be able to obtain height.

### Simulation of covariates

Covariates will be sampled from the following distributions :

-   CRCL : Normal distribution with a minimum of 10 mL/min, with mean
    `r signif(carlier_crcl$norm_par["mean"],3)` mL/min and standard
    deviation `r signif(carlier_crcl$norm_par["sd"],3)` mL/min
-   WT : Lognormal distribution with mean
    `r signif(exp(carlier_wt$lnorm_par["meanlog"]),3)` kg, and
    coefficient of variation
    `r signif(compute_cv_from_sd_lnorm(carlier_wt$lnorm_par["sdlog"]), 3)*100`
    %
-   AGE : Normal distribution with mean
    `r signif(carlier_age$norm_par["mean"],3)` mL/min and standard
    deviation `r signif(carlier_age$norm_par["sd"],3)` years and with a
    minimum of 18 years
-   BMI : Log-normal distribution with mean
    `r signif(carlier_bmi$lnorm_par["meanlog"],3)` mL/min and standard
    deviation `r signif(carlier_bmi$lnorm_par["sdlog"],3)` kg/m\^{2}

```{r functions-sim-cov, warning=FALSE, message=FALSE}
sim_norm_without_negatives <- function(n,mean,sd){
      
  #' Simulate values from normal distribution but resample negative values
  #'
  #' @param n numeric. number of samples to simulate
  #' @param mean numeric. mean of the normal distribution
  #' @param sd numeric. standard deviation of the normal distribtuion
  #' @return numeric vector of n values
  sim_values <- rnorm(n,mean,sd)
  neg_values <- sim_values[sim_values<=0]
  pos_values <- sim_values[sim_values>0]

  while(length(neg_values)>0){
    new_values <- rnorm(length(neg_values),mean,sd)
    sim_values <- c(pos_values,new_values)
    neg_values <- sim_values[sim_values<=0]
    pos_values <- sim_values[sim_values>0]
    }
  sim_values
}

# Only for normal distribution when the minimum is 10 ml/min, resample below 10
sim_norm_without_dialysis <- function(n,mean,sd){
      
  #' Simulate values from normal distribution but resample negative values
  #'
  #' @param n numeric. number of samples to simulate
  #' @param mean numeric. mean of the normal distribution
  #' @param sd numeric. standard deviation of the normal distribtuion
  #' @return numeric vector of n values
  sim_values <- rnorm(n,mean,sd)
  neg_values <- sim_values[sim_values<=10]
  pos_values <- sim_values[sim_values>10]

  while(length(neg_values)>0){
    new_values <- rnorm(length(neg_values),mean,sd)
    sim_values <- c(pos_values,new_values)
    neg_values <- sim_values[sim_values<=10]
    pos_values <- sim_values[sim_values>10]
    }
  sim_values
}
# Resimulate subjects younger than 18 years for normally distributed age
sim_norm_without_minors <- function(n,mean,sd){
      
  #' Simulate values from normal distribution but resample negative values
  #'
  #' @param n numeric. number of samples to simulate
  #' @param mean numeric. mean of the normal distribution
  #' @param sd numeric. standard deviation of the normal distribtuion
  #' @return numeric vector of n values
  sim_values <- rnorm(n,mean,sd)
  neg_values <- sim_values[sim_values<18]
  pos_values <- sim_values[sim_values>=18]

  while(length(neg_values)>0){
    new_values <- rnorm(length(neg_values),mean,sd)
    sim_values <- c(pos_values,new_values)
    neg_values <- sim_values[sim_values<18]
    pos_values <- sim_values[sim_values>=18]
    }
  sim_values
}
```

The relevant part of the correlation matrix is converted to a covariance
matrix by multiplying the already calculated standard deviations of the
variables with the diagonal of the correlation matrix. More information
can be found at
[SAS](https://blogs.sas.com/content/iml/2010/12/10/converting-between-correlation-and-covariance-matrices.html)
and at
[stats](https://stats.stackexchange.com/questions/62850/obtaining-covariance-matrix-from-correlation-matrix).
A multivariate normal distribution is fitted with resimulation of CRCL
below 10 ml/min, AGE below 18 years, and other variables below 0. 

The proportion of males is 0.85 in the original article, so the number of males is 531 and the number of females is 94 to add up to 625.

```{r sim-cov-carlier, warning=FALSE, message=FALSE}
mean_WT_log <- as.numeric(carlier_wt$lnorm_par["meanlog"])
sd_WT_log <- as.numeric(carlier_wt$lnorm_par["sdlog"])
mean_BMI_log <- as.numeric(carlier_bmi$lnorm_par["meanlog"])
sd_BMI_log <- as.numeric(carlier_bmi$lnorm_par["sdlog"])
r_M <- 0.34 # Pearson correlation between HT_log and WT_log for males
r_F <- 0.36 # Pearson correlation for females

# sd is different based on sex
mean_HT_log = (mean_WT_log - mean_BMI_log) / 2
sd_HT_M_log = (-r_M * sd_WT_log + sqrt(r_M^2 * sd_WT_log^2 - sd_WT_log^2 + sd_BMI_log^2))/2 # quadratic equation with positive result
sd_HT_F_log = (-r_F * sd_WT_log + sqrt(r_F^2 * sd_WT_log^2 - sd_WT_log^2 + sd_BMI_log^2))/2 # quadratic equation with positive

correlated_simulation <- function(n, means, cov_matrix) {
  set.seed(1991)
  collected <- matrix(NA, 0, length(means))
  colnames(collected) <- names(means)

  while (nrow(collected) < n) {
    batch <- MASS::mvrnorm(n = n, mu = means, Sigma = cov_matrix)
    colnames(batch) <- names(means)

    valid <- batch[, "AGE"] >= 18 &
             batch[, "CRCL_MDRD"] >= 10

    batch_valid <- batch[valid, , drop = FALSE]
    collected <- rbind(collected, batch_valid)
  }

  collected[1:n, , drop = FALSE]
}

means <- c(
  CRCL_MDRD = carlier_crcl$norm_par["mean"],
  WT_log      = carlier_wt$lnorm_par["meanlog"],
  AGE     = carlier_age$norm_par["mean"],
  HT_log     = mean_HT_log
)

sds_M <- c(
  CRCL_MDRD = carlier_crcl$norm_par["sd"],
  WT_log      = carlier_wt$lnorm_par["sdlog"], 
  AGE     = carlier_age$norm_par["sd"],
  HT_log     = sd_HT_M_log
)

sds_F <- c(
  CRCL_MDRD = carlier_crcl$norm_par["sd"],
  WT_log      = carlier_wt$lnorm_par["sdlog"], 
  AGE     = carlier_age$norm_par["sd"],
  HT_log     = sd_HT_F_log
)

covariates <- c("CRCL_MDRD", "WT_log", "AGE", "HT_log")

names(means) <- covariates

# Correlation and Covariance Matrix
cor_carlier_M <- cor_matrix_Carlier_M[covariates, covariates]
cov_matrix_M <- diag(sds_M) %*% cor_carlier_M %*% diag(sds_M)
eigen(cov_matrix_M)$values

cor_carlier_F <- cor_matrix_Carlier_F[covariates, covariates]
cov_matrix_F <- diag(sds_F) %*% cor_carlier_F %*% diag(sds_F)
eigen(cov_matrix_F)$values

# Simulate
sim_data_M <- correlated_simulation(n = 531, means = means, cov_matrix = cov_matrix_M)
sim_data_M <- as_tibble(sim_data_M) %>%
  mutate(SEX = 0)

sim_data_F <- correlated_simulation(n = 94, means = means, cov_matrix = cov_matrix_F)
sim_data_F <- as_tibble(sim_data_F) %>%
  mutate(SEX = 1)

sim_data <- rbind(sim_data_M, sim_data_F)

# Put the covariates together
sim_cov_carlier <- tibble(
  Paper = "Carlier_2013",
  ID_within_paper = 1:n_patient,
  ICU = 1,
  BURN = 0,
  OBESE = 0,
  CRCL = sim_data$CRCL_MDRD,
  WT_log   = sim_data$WT_log,
  AGE  = sim_data$AGE,
  HT_log  = sim_data$HT_log,
  SEX = sim_data$SEX
)

# Add OBESE variable and reconvert logWT to WT
sim_cov_carlier <- sim_cov_carlier %>%
  mutate(WT = exp(WT_log),
         HT = (exp(HT_log))*100,
         BMI = WT / (HT/100)^2,
         OBESE = ifelse(BMI > 30, 1, 0))

summary_sim_cov_carlier <- sim_cov_carlier |> 
  pivot_longer(cols = c(ICU,BURN,OBESE,CRCL,WT,AGE,BMI),
               names_to = "Covariate") |> 
  summarise(.by = c(Covariate,Paper),
            Min = min(value),
            Q1 = quantile(value, 0.25),
            Median = quantile(value, 0.5),
            Q3 = quantile(value, 0.75),
            Max = max(value))
```

```{r carlier-quantile-table, warning=FALSE, message=FALSE}
cov_carlier_compare <- summary_sim_cov_carlier |>
  rename_with(~ paste0(.x, "_sim"), -one_of("Covariate", "Paper")) |>
  left_join(rename_with(
    cov_data_carlier,
    ~ paste0(.x, "_true"),
    -one_of("Covariate", "Paper", "Unit")
  )) |> 
  relocate(Covariate, .after = Paper) |> 
  relocate(Unit, .after = Covariate)

cov_carlier_compare |>
  gt() |>
  fmt_scientific() |> 
    tab_spanner(columns = starts_with("Min"),
              label = "Min") |> 
      tab_spanner(columns = starts_with("Q1"),
              label = "Q1") |> 
      tab_spanner(columns = starts_with("Median"),
              label = "Median") |> 
      tab_spanner(columns = starts_with("Q3"),
              label = "Q3") |> 
  tab_spanner(columns = starts_with("Max"),
              label = "Max")


```

```{r function-plot-cov, warning=FALSE, message=FALSE}
evaluate_cov_sim <- function(p, q, cov_name, cov_unit, cov_sim) {
  #' Plot simulated covariate distribution versus observed quantiles from the paper
  #' @param p numeric vector of probabilities
  #' @param q numeric vector of quantiles
  #' @param cov_name character string containing the covariate name
  #' @param cov_unit character string containing the unit of the covariate
  #' @param cov_sim tibble containing the simulated values for the covariate
  #' @return ggplot2 showing the distributions of the simulated covariate and the observed covariate quantiles
  
  
  #create a dataset with the true percentiles to be able to show them on the plot
  true_percentiles <- tibble(p, q) |> 
    mutate(label = case_match(p, 
                              0 ~ "True min",
                              0.25 ~ "True Q1",
                              0.5 ~ "True median",
                              0.75 ~ "True Q3",
                              1 ~ "True max")) 
  
  #create a dataset with the percentiles of the fitted distributions to be able to show them on the plot
   sim_percentiles <-  tibble(
    probability = p,
    q = quantile(pull(cov_sim,cov_name),p) 
  ) |> 
        mutate(label = case_match(probability, 
                              0 ~ "Sim min",
                              0.25 ~ "Sim Q1",
                              0.5 ~ "Sim median",
                              0.75 ~ "Sim Q3",
                              1 ~ "Sim max"))
   
# create an initial histogram in order to be able to extract statistics of the binned covariate
hist_ini <- ggplot(cov_sim, aes(x = .data[[cov_name]])) +
  geom_histogram(bins=15)
   
hist_data <- layer_data(hist_ini)
   
plot <- ggplot(hist_data) +
  geom_rect(aes(xmin=xmin,xmax=xmax, ymin=0, ymax=count),
            alpha = 0.5) +
     geom_vline(data = true_percentiles,
                mapping = aes(xintercept = q,color = "True"),
                lty = "dashed") +
     geom_vline(data = sim_percentiles,
                mapping = aes(xintercept = q,color = "Sim"),
                lty = "solid") +
     theme_bw() +
     ylab("Count") +
     xlab(paste0(cov_name, " (", cov_unit, ")"))  +


     geom_text(
       data = true_percentiles,
       mapping = aes(x = q, label = label, color = "True"),
       show.legend = FALSE,
       y = 0.95*max(hist_data$ymax),
       hjust = 1,
       vjust = -1,
       angle = 90
     ) +
     geom_text(
       data = sim_percentiles,
       mapping = aes(x = q, label = label, color = "Sim"),
       show.legend = FALSE,
       y = 0.95*max(hist_data$ymax),
       hjust = 1,
       vjust = -1,
       angle = 90
     ) +
     scale_color_brewer(name = "Quantiles",
                        palette ="Dark2")
   plot
}

plot_sim_cov_wrapper <- function(cov_data,cov_name,cov_sim){
  cov_df <- extract_cov_param(cov_data, cov_name)

evaluate_cov_sim(cov_df$p,cov_df$q,cov_df$cov_name,cov_df$cov_unit,
                 cov_sim)
}
```

```{r plot-cov-carlier, warning=FALSE, message=FALSE}
plot_sim_cov_wrapper(cov_data_carlier,"CRCL",sim_cov_carlier)
plot_sim_cov_wrapper(cov_data_carlier,"WT",sim_cov_carlier)
plot_sim_cov_wrapper(cov_data_carlier,"AGE",sim_cov_carlier)
plot_sim_cov_wrapper(cov_data_carlier,"BMI",sim_cov_carlier)
```

Then, based on BMI and WT, height (HT) is calculated to be able to
calculate the body surface area (BSA). CRCL was calculated using the
measured urinary equation, and as we have know information about the
urinary volume or urinary creatinine concentration, we cannot reconvert
CRCL to CREAT based on this equation. The most recommended equation is
CKD-EPI, however as different exponents are used based on the value of
CREAT, it can't be used for reconversion of CRCL to CREAT. Thus, the
second best equation, MDRD is used.

```{r reconvert_creat_carlier, warning = FALSE, message = FALSE}
reconvert_MDRD <- function(AGE, CRCL, SEX) {
    # Mutliply by 0.742 for females
    sex_factor <- ifelse(SEX == 0, 1, 0.742)
    
    # Calculate DFG
   CREAT <- (CRCL / (175 * (AGE^(-0.203)) * sex_factor))^(-1 / 1.154)
  
  return(CREAT)
}

sim_cov_carlier <- sim_cov_carlier %>%
  mutate(
    BSA = (WT^0.425 * (HT^0.725) * 0.007184), # Calculate body surface area based on Dubois&Dubois formula
    CRCL = CRCL * (1.73 /BSA) # convert mL/min to mL/min/1.73 m2
  )
sim_cov_carlier$CREAT <- mapply(reconvert_MDRD, sim_cov_carlier$AGE, sim_cov_carlier$CRCL, sim_cov_carlier$SEX)
sim_cov_carlier <- sim_cov_carlier %>%
  dplyr::select(Paper, ID_within_paper, ICU, BURN, OBESE, CREAT, WT, BSA, BMI, AGE, SEX, HT)

# Calculate the simulated correlation matrices separately for the two sexes to compare with the original
sim_cov_carlier_M <- sim_cov_carlier %>% filter(SEX == 0)
sim_cov_carlier_F <- sim_cov_carlier %>% filter(SEX == 1)

cor_sim_M <- cor(sim_cov_carlier_M %>% dplyr::select(WT, CREAT, AGE, HT, BMI), use = "complete.obs")
cor_sim_F <- cor(sim_cov_carlier_F %>% dplyr::select(WT, CREAT, AGE, HT, BMI), use = "complete.obs")

ggcorrplot(cor_matrix_Carlier_F, lab = TRUE, title = "Carlier original (females)")
Carlier_F <- ggcorrplot(cor_sim_F, lab = TRUE, title = "Carlier simulated (females)")
Carlier_F 
ggcorrplot(cor_matrix_Carlier_M, lab = TRUE, title = "Carlier original (males)")
Carlier_M <- ggcorrplot(cor_sim_M, lab = TRUE, title = "Carlier simulated (males)")
Carlier_M
```

## @fournier2018

## General considerations about the paper

This study includes only 21 patients, thus deriving distributions from
the reported covariates values might be difficult

### Covariate distribution

All patients are ICU patients with burns who are assumed to be non
obese. Creatinine clearance is calculated based on the Cockcroft & Gault
equation, whose output is in mL/min, so there is no need to calculate
BSA. Thus for all patients ICU = 1, BURN = 1 and OBESE=0.

It is not mentioned in the article if certain patients are on renal
replacement therapy, therefore we assume that they are not, and set the
minimum value of creatinine clearance to 10 mL/min. CRCL was estimated
using the Cockroft and Gault equation.

For AGE, not the median and IQR is given but the mean (50.1) and
standard deviation (24.3). As we cannot compare the simulated
distribution to the true one, so we suppose that the distribution
follows a normal distribution.

The proportion of males in the original article is 0.762.

Neither BMI nor HT or BSA are provided in the article, thus the height
is simulated based on the height obtained from the MIMIC clinical
dataset (normal distribution with mean of 169 and sd of 10.3).

For WT, AGE and CRCL here are the data from the paper :

```{r fournier-cov-data-paper, warning=FALSE, message=FALSE}
cov_data_fournier <- cov_data |> 
  filter(Paper == "Fournier_2018")

cov_data_fournier |> 
  filter(Covariate %in% c("WT","CRCL","AGE")) |> 
  kable()
```

```{r Fournier-cor-matrix, warning=FALSE, message=FALSE}
cor_matrix_Fournier_F <- readRDS(here("Simulations/cor_matrix_Fournier_F.rds"))
print(cor_matrix_Fournier_F)
ggcorrplot(cor_matrix_Fournier_F, lab = TRUE, title = "Sex: female")

cor_matrix_Fournier_M <- readRDS(here("Simulations/cor_matrix_Fournier_M.rds"))
print(cor_matrix_Fournier_M)
ggcorrplot(cor_matrix_Fournier_M, lab = TRUE, title = "Sex: male")
```

```{r fit-fournier-crcl, warning=FALSE, message=FALSE}
#| warning: false

fournier_crcl <- fit_cov_wrapper(
  cov_data = cov_data_fournier,
  cov_name = "CRCL",
  min_plot = 0,
  max_plot = 250,
  tol=0.002
)

fournier_crcl$plot
```

Tolerance was increased to 0.002 to have a fit for log-normal. Normal
distribution is better suited.

```{r fit-fournier-wt, warning=FALSE, message=FALSE}
#| warning: false
fournier_wt <- fit_cov_wrapper(
  cov_data = cov_data_fournier,
  cov_name = "WT",
  min_plot = 40,
  max_plot = 120
)

fournier_wt$plot
```

In this case the log-normal distribution quantiles are closer to the
true quantiles than the normal ones. We will take the log-normal ones.

### Simulation of covariates

Covariates will be sampled from the following distributions :

-   CRCL : Normal distribution with mean
    `r signif(fournier_crcl$norm_par["mean"])` mL/min and standard
    deviation `r signif(fournier_crcl$norm_par["sd"],3)` mL/min
-   WT : Lognormal distribution with mean
    `r signif(exp(fournier_wt$lnorm_par["meanlog"]),3)` kg, and
    coefficient of variation
    `r signif(compute_cv_from_sd_lnorm(fournier_wt$lnorm_par["sdlog"]), 3) * 100`
    %
-   AGE : Supposed normal distribution with mean of 50.1 years and
    standard deviation of 24.3

The proportion of males is 0.762 in the original article, so the number of males is 476 and the number of females is 149 to add up to 625.

```{r sim-cov-fournier, warning=FALSE, message=FALSE}
correlated_simulation <- function(n, means, cov_matrix) {
  set.seed(1991)
  collected <- matrix(NA, 0, length(means))
  colnames(collected) <- names(means)

  while (nrow(collected) < n) {
    batch <- MASS::mvrnorm(n = n, mu = means, Sigma = cov_matrix)
    colnames(batch) <- names(means)

    valid <- batch[, "AGE"] >= 18 &
             batch[, "CRCL_CG"] >= 10

    batch_valid <- batch[valid, , drop = FALSE]
    collected <- rbind(collected, batch_valid)
  }

  collected[1:n, , drop = FALSE]
}

means <- c(
  CRCL_CG = fournier_crcl$norm_par["mean"],
  WT_log      = fournier_wt$lnorm_par["meanlog"],
  AGE     = 50.1,
  HT = 169
)

sds <- c(
  CRCL_CG = fournier_crcl$norm_par["sd"],
  WT_log      = fournier_wt$lnorm_par["sdlog"],
  AGE     = 24.3,
  HT = 10.3
)

covariates <- c("CRCL_CG", "WT_log", "AGE", "HT")

names(means) <- covariates

# Correlation and Covariance Matrix
cor_fournier_M <- cor_matrix_Fournier_M[covariates, covariates]
cov_matrix_M <- diag(sds) %*% cor_fournier_M %*% diag(sds)
eigen(cov_matrix_M)$values

cor_fournier_F <- cor_matrix_Fournier_F[covariates, covariates]
cov_matrix_F <- diag(sds) %*% cor_fournier_F %*% diag(sds)
eigen(cov_matrix_F)$values

# Simulate
sim_data_M <- correlated_simulation(n = 476, means = means, cov_matrix = cov_matrix_M)
sim_data_M <- as_tibble(sim_data_M) %>%
  mutate(SEX = 0)

sim_data_F <- correlated_simulation(n = 149, means = means, cov_matrix = cov_matrix_F)
sim_data_F <- as_tibble(sim_data_F) %>%
  mutate(SEX = 1)

sim_data <- rbind(sim_data_M, sim_data_F)

# Put the covariates together
sim_cov_fournier <- tibble(
  Paper = "Fournier_2018",
  ID_within_paper = 1:n_patient,
  ICU = 1,
  BURN = 1,
  OBESE = 0,
  CRCL = sim_data$CRCL_CG,
  WT_log   = sim_data$WT_log,
  AGE  = sim_data$AGE,
  HT  = sim_data$HT,
  SEX = sim_data$SEX)

sim_cov_fournier <- sim_cov_fournier %>%
  mutate(WT = exp(WT_log))

summary_sim_cov_fournier <- sim_cov_fournier |> 
  pivot_longer(cols = c(ICU,BURN,OBESE,CRCL,WT,AGE),
               names_to = "Covariate") |> 
  summarise(.by = c(Covariate,Paper),
            Min = min(value),
            Q1 = quantile(value, 0.25),
            Median = quantile(value, 0.5),
            Q3 = quantile(value, 0.75),
            Max = max(value))

paste0("The proportion of males is ", round(mean(sim_cov_fournier$SEX == 0), 3), " compared to the original 0.762.")
```

```{r fournier-quantile-table, warning=FALSE, message=FALSE}
cov_fournier_compare <- summary_sim_cov_fournier |>
  rename_with(~ paste0(.x, "_sim"), -one_of("Covariate", "Paper")) |>
  left_join(rename_with(
    cov_data_fournier,
    ~ paste0(.x, "_true"),
    -one_of("Covariate", "Paper", "Unit")
  )) |> 
  relocate(Covariate, .after = Paper) |> 
  relocate(Unit, .after = Covariate)

cov_fournier_compare |>
  gt() |>
  fmt_scientific() |> 
    tab_spanner(columns = starts_with("Min"),
              label = "Min") |> 
      tab_spanner(columns = starts_with("Q1"),
              label = "Q1") |> 
      tab_spanner(columns = starts_with("Median"),
              label = "Median") |> 
      tab_spanner(columns = starts_with("Q3"),
              label = "Q3") |> 
  tab_spanner(columns = starts_with("Max"),
              label = "Max")
```

```{r plot-cov-fournier, warning=FALSE, message=FALSE}
plot_sim_cov_wrapper(cov_data_fournier,"CRCL",sim_cov_fournier)
plot_sim_cov_wrapper(cov_data_fournier,"WT",sim_cov_fournier)
plot_sim_cov_wrapper(cov_data_fournier,"AGE",sim_cov_fournier)
```

Then, based on the BMI and WT, the height (HT) is calculated to be able
to calculate the body surface area (BSA). CRCL is reconverted to serum
creatinine (CREAT) based on the Cockroft and Gault equation.

```{r reconvert_creat_fournier, warning = FALSE, message = FALSE}
reconvert_CG <- function(AGE, CRCL, SEX, WT) {
    # Mutliply by 0.85 for females
    sex_factor <- ifelse(SEX == 0, 1, 0.85)
    
    # Calculate DFG
    CREAT = ((140 - AGE) * WT * sex_factor) / (CRCL * 72)
  
  return(CREAT)
}

sim_cov_fournier$CREAT <- mapply(reconvert_CG, sim_cov_fournier$AGE, sim_cov_fournier$CRCL, sim_cov_fournier$SEX, sim_cov_fournier$WT)
sim_cov_fournier <- sim_cov_fournier %>%
  mutate(
     BSA = (WT^0.425 * (HT^0.725) * 0.007184),
     BMI = WT / (HT/100)^2,
     OBESE = ifelse(BMI > 30, 1, 0)
  ) %>%
  dplyr::select(Paper, ID_within_paper, ICU, BURN, OBESE, CREAT, WT, BSA, BMI, AGE, SEX, HT)
# Calculate the simulated correlation matrices separately for the two sexes to compare with the original
sim_cov_fournier_M <- sim_cov_fournier %>% filter(SEX == 0)
sim_cov_fournier_F <- sim_cov_fournier %>% filter(SEX == 1)

cor_sim_M <- cor(sim_cov_fournier_M %>% dplyr::select(WT, CREAT, AGE, HT, BMI), use = "complete.obs")
cor_sim_F <- cor(sim_cov_fournier_F %>% dplyr::select(WT, CREAT, AGE, HT, BMI), use = "complete.obs")

ggcorrplot(cor_matrix_Fournier_F, lab = TRUE, title = "Fournier original (females)")
Fournier_F <- ggcorrplot(cor_sim_F, lab = TRUE, title = "Fournier simulated (females)")
Fournier_F
ggcorrplot(cor_matrix_Fournier_M, lab = TRUE, title = "Fournier original (males)")
Fournier_M <- ggcorrplot(cor_sim_M, lab = TRUE, title = "Fournier simulated (males)")
Fournier_M
```

## @mellon2020

### Covariate distribution

Patients are obese, but otherwise healthy, meaning that they are not in
intensive care and not burn victims. They are all obese (BMI \> 30
kg/m\^2). Creatinine clearance is calculated based on the MDRD equation.
Thus for all patients ICU = 0, BURN = 0, OBESE=1.

For WT, BMI and CRCL here are the data from the paper :

```{r mellon-cov-data-paper, warning=FALSE, message=FALSE}
cov_data_mellon <- cov_data |> 
  filter(Paper == "Mellon_2020")

cov_data_mellon |> 
  filter(Covariate %in% c("WT","CRCL", "AGE", "BMI", "BMI_log")) |> 
  kable()
```

```{r Mellon-cor-matrix, warning=FALSE, message=FALSE}
cor_matrix_Mellon_F <- readRDS(here("Simulations/cor_matrix_Mellon_F.rds"))
print(cor_matrix_Mellon_F)
ggcorrplot(cor_matrix_Mellon_F, lab = TRUE, title = "Sex: female")

cor_matrix_Mellon_M <- readRDS(here("Simulations/cor_matrix_Mellon_M.rds"))
print(cor_matrix_Mellon_M)
ggcorrplot(cor_matrix_Mellon_M, lab = TRUE, title = "Sex: male")
```

Mellon is different from the other papers as CRCL is given as
ml/min/1.73m\^{2} based on the MDRD equation, and instead of Q1 and Q3,
the minimum and maxiumum of the covariates is given (apart from BMI).
Another particularity is that due to the particularity of the subjects
(obesity), the distribution of body weight is positively (right) skewed.
Positively skewed distribution is relatively rare, so it is not easy to
find a distribution function to describe it. Beta distribution needs
scaling the data between 0 and 1. and 4 parameter stable distribution et
gamma distribution have several parameters that need to be defined, and
in the absence of the original data, it is not possible to estimate
these parameters.

The proportion of males is 0.148.

```{r fit-mellon-crcl, warning=FALSE, message=FALSE}
#| warning: false

mellon_crcl <- fit_cov_wrapper(
  cov_data = cov_data_mellon,
  cov_name = "CRCL",
  min_plot = 0,
  max_plot = 250
)

mellon_crcl$plot
```

Similar fits; normal distribution is selected.

```{r fit-mellon-wt, warning=FALSE, message=FALSE}
#| warning: false
mellon_wt <- fit_cov_wrapper(
  cov_data = cov_data_mellon,
  cov_name = "WT",
  min_plot = 40,
  max_plot = 170
)

mellon_wt$plot
```

Log-normal distribution fits slightly better than normal.

```{r fit-mellon-age, warning=FALSE, message=FALSE}
#| warning: false
mellon_age <- fit_cov_wrapper(
  cov_data = cov_data_mellon,
  cov_name = "AGE",
  min_plot = 0,
  max_plot = 120
)

mellon_age$plot
```

Normal distribution fits a bit better.

```{r fit-mellon-bmi, warning=FALSE, message=FALSE}
#| warning: false
mellon_bmi <- fit_cov_wrapper(
  cov_data = cov_data_mellon,
  cov_name = "BMI",
  min_plot = 0,
  max_plot = 100, 
  tol = 0.001
)

mellon_bmi$plot
```

```{r fit-mellon-bmi-log, warning=FALSE, message=FALSE}
#| warning: false
mellon_bmi_log <- fit_cov_wrapper(
  cov_data = cov_data_mellon,
  cov_name = "BMI_log",
  min_plot = 2.5,
  max_plot = 5, 
  tol = 0.001
)

mellon_bmi_log$plot
```

Log-normal distribution fits a bit better.

### Simulation of covariates

Covariates will be sampled from the following distributions :

-   CRCL : Normal distribution from which negative values have been
    resampled, with mean `r signif(mellon_crcl$norm_par["mean"],3)`
    mL/min and standard deviation
    `r signif(mellon_crcl$norm_par["sd"],3)` mL/min
-   WT : Log-normal distribution with mean
    `r signif(exp(mellon_wt$lnorm_par["meanlog"]),3)` kg, and
    coefficient of variation
    `r signif(compute_cv_from_sd_lnorm(mellon_wt$lnorm_par["sdlog"]), 3)*100`
    %
-   AGE : Normal distribution with mean
    `r signif(exp(mellon_age$norm_par["mean"]),3)` years, and standard
    deviation `r signif(mellon_age$norm_par["sd"],3)` %
-   BMI : Log-normal distribution with mean
    `r signif(exp(mellon_bmi$lnorm_par["meanlog"]),3)` $kg/m^{2}$, and
    coefficient of variation
    `r signif(compute_cv_from_sd_lnorm(mellon_bmi$lnorm_par["sdlog"]), 3)*100`
    %
    
The proportion of males is 0.148 in the original article, so the number of males is 93 and the number of females is 532 to add up to 625.

```{r sim-cov-mellon, warning=FALSE, message=FALSE}
correlated_simulation <- function(n, means, cov_matrix) {
  set.seed(1991)
  collected <- matrix(NA, 0, length(means))
  colnames(collected) <- names(means)

  while (nrow(collected) < n) {
    batch <- MASS::mvrnorm(n = n, mu = means, Sigma = cov_matrix)
    colnames(batch) <- names(means)

    valid <- batch[, "AGE"] >= 18 &
             batch[, "CRCL_MDRD_norm"] >= 10 &
             batch[, "BMI_log"] > 3.401197 # log(30)

    batch_valid <- batch[valid, , drop = FALSE]
    collected <- rbind(collected, batch_valid)
  }

  collected[1:n, , drop = FALSE]
}

means <- c(
  CRCL_MDRD_norm = mellon_crcl$norm_par["mean"],
  WT_log      = mellon_wt$lnorm_par["meanlog"],
  AGE     = mellon_age$norm_par["mean"],
  BMI_log     = mellon_bmi$lnorm_par["meanlog"]
)

sds <- c(
  CRCL_MDRD_norm = mellon_crcl$norm_par["sd"],
  WT_log      = mellon_wt$lnorm_par["sdlog"], 
  AGE = mellon_age$norm_par["sd"],
  BMI_log     = mellon_bmi$lnorm_par["sdlog"]
)

covariates <- c("CRCL_MDRD_norm", "WT_log", "AGE", "BMI_log")

names(means) <- covariates

# Correlation and Covariance Matrix
cor_mellon_M <- cor_matrix_Mellon_M[covariates, covariates]
cov_matrix_M <- diag(sds) %*% cor_mellon_M %*% diag(sds)
eigen(cov_matrix_M)$values

cor_mellon_F <- cor_matrix_Mellon_F[covariates, covariates]
cov_matrix_F <- diag(sds) %*% cor_mellon_F %*% diag(sds)
eigen(cov_matrix_F)$values

# Simulate
sim_data_M <- correlated_simulation(n = 93, means = means, cov_matrix = cov_matrix_M)
sim_data_M <- as_tibble(sim_data_M) %>%
  mutate(SEX = 0)

sim_data_F <- correlated_simulation(n = 532, means = means, cov_matrix = cov_matrix_F)
sim_data_F <- as_tibble(sim_data_F) %>%
  mutate(SEX = 1)

sim_data <- rbind(sim_data_M, sim_data_F)

# Put the covariates together
sim_cov_mellon <- tibble(
  Paper = "Mellon_2020",
  ID_within_paper = 1:n_patient,
  ICU = 0,
  BURN = 0,
  OBESE = 1,
  CRCL = sim_data$CRCL_MDRD_norm,
  WT_log   = sim_data$WT_log,
  AGE  = sim_data$AGE,
  BMI_log  = sim_data$BMI_log,
  SEX = sim_data$SEX)

sim_cov_mellon <- sim_cov_mellon %>%
  mutate(WT = exp(WT_log),
         BMI = exp(BMI_log))

summary_sim_cov_mellon <- sim_cov_mellon |> 
  pivot_longer(cols = c(ICU,BURN,OBESE,CRCL,WT,AGE),
               names_to = "Covariate") |> 
  summarise(.by = c(Covariate,Paper),
            Min = min(value),
            Q1 = quantile(value, 0.25),
            Median = quantile(value, 0.5),
            Q3 = quantile(value, 0.75),
            Max = max(value))
```

```{r mellon-quantile-table, warning=FALSE, message=FALSE}
sim_cov_mellon <- sim_cov_mellon %>%
  dplyr::select(Paper, ID_within_paper, ICU, BURN, OBESE, CRCL, WT, AGE, BMI, SEX)

cov_mellon_compare <- summary_sim_cov_mellon |>
  rename_with(~ paste0(.x, "_sim"), -one_of("Covariate", "Paper")) |>
  left_join(rename_with(
    cov_data_mellon,
    ~ paste0(.x, "_true"),
    -one_of("Covariate", "Paper", "Unit")
  )) |> 
  relocate(Covariate, .after = Paper) |> 
  relocate(Unit, .after = Covariate)

cov_mellon_compare |>
  gt() |>
  fmt_scientific() |> 
  tab_spanner(columns = starts_with("Min"),
              label = "Min") |> 
  tab_spanner(columns = starts_with("Q1"),
              label = "Q1") |> 
  tab_spanner(columns = starts_with("Median"),
              label = "Median") |> 
  tab_spanner(columns = starts_with("Q3"),
              label = "Q3") |> 
  tab_spanner(columns = starts_with("Max"),
              label = "Max")
```

Then, based on the BMI and WT, the height (HT) is calculated to be able
to calculate the body surface area (BSA). CRCL is reconverted to serum
creatinine (CREAT) based on the MDRD equation.

```{r reconvert_creat_mellon, warning = FALSE, message = FALSE}
reconvert_MDRD <- function(AGE, CRCL, SEX) {
    # Mutliply by 0.85 for females
    sex_factor <- ifelse(SEX == 0, 1, 0.742)
    
    # Calculate DFG
   CREAT <- (CRCL / (175 * (AGE^(-0.203)) * sex_factor))^(-1 / 1.154)
  
  return(CREAT)
}

sim_cov_mellon$CREAT <- mapply(reconvert_MDRD, sim_cov_mellon$AGE, sim_cov_mellon$CRCL, sim_cov_mellon$SEX)
sim_cov_mellon <- sim_cov_mellon %>%
  mutate(
   HT = sqrt(WT/BMI) * 100, # mutliplied by 100 to convert m to cm
    BSA = (WT^0.425 * (HT^0.725) * 0.007184),
    OBESE = ifelse(BMI > 30, 1, 0)
  ) %>%
  dplyr::select(Paper, ID_within_paper, ICU, BURN, OBESE, CREAT, WT, BSA, BMI, AGE, SEX, HT)

# Calculate the simulated correlation matrices separately for the two sexes to compare with the original
sim_cov_mellon_M <- sim_cov_mellon %>% filter(SEX == 0)
sim_cov_mellon_F <- sim_cov_mellon %>% filter(SEX == 1)

cor_sim_M <- cor(sim_cov_mellon_M %>% dplyr::select(WT, CREAT, AGE, HT, BMI), use = "complete.obs")
cor_sim_F <- cor(sim_cov_mellon_F %>% dplyr::select(WT, CREAT, AGE, HT, BMI), use = "complete.obs")

ggcorrplot(cor_matrix_Mellon_F, lab = TRUE, title = "Mellon original (females)")
Mellon_F <- ggcorrplot(cor_sim_F, lab = TRUE, title = "Mellon simulated (females)")
Mellon_F
ggcorrplot(cor_matrix_Mellon_M, lab = TRUE, title = "Mellon original (males)")
Mellon_M <- ggcorrplot(cor_sim_M, lab = TRUE, title = "Mellon simulated (males)")
Mellon_M
```

## @rambaud2020

### Covariate distribution

All patients are ICU patients with endocarditis. They are not considered
to be burn victims or obese. Creatinine clearance was calculated using
the CKD-EPI equation. Thus for all patients ICU = 1, BURN = 0 and
OBESE=0. In this paper, the distribution of serum creatinine is given
(in micromol/L), so it can be directly simulated instead of reconverting
CRCL. The proportion of males is 0.794.

For WT and CRCL here are the data from the paper :

```{r rambaud-cov-data-paper, warning=FALSE, message=FALSE}
cov_data_rambaud <- cov_data |> 
  filter(Paper == "Rambaud_2020")

cov_data_rambaud |> 
  filter(Covariate %in% c("WT","CREAT","AGE","HT")) |> 
  kable()
```

```{r Rambaud-cor-matrix, warning=FALSE, message=FALSE}
cor_matrix_Rambaud_F <- readRDS(here("Simulations/cor_matrix_Rambaud_F.rds"))
print(cor_matrix_Rambaud_F)
ggcorrplot(cor_matrix_Rambaud_F, lab = TRUE, title = "Sex: female")

cor_matrix_Rambaud_M <- readRDS(here("Simulations/cor_matrix_Rambaud_M.rds"))
print(cor_matrix_Rambaud_M)
ggcorrplot(cor_matrix_Rambaud_M, lab = TRUE, title = "Sex: male")
```

```{r fit-rambaud-creat, warning=FALSE, message=FALSE}
#| warning: false

rambaud_creat <- fit_cov_wrapper(
  cov_data = cov_data_rambaud,
  cov_name = "CREAT",
  min_plot = 0,
  max_plot = 250
)

rambaud_creat$plot

```

Log-normal distribution fits the best the data, so it is selected.

```{r fit-rambaud-wt, warning=FALSE, message=FALSE}
#| warning: false
rambaud_wt <- fit_cov_wrapper(
  cov_data = cov_data_rambaud,
  cov_name = "WT",
  min_plot = 40,
  max_plot = 120
)

rambaud_wt$plot
```

No discernable difference between the two fits. The log-normal distribution
is selected.

```{r fit-rambaud-age, warning=FALSE, message=FALSE}
#| warning: false
rambaud_age <- fit_cov_wrapper(
  cov_data = cov_data_rambaud,
  cov_name = "AGE",
  min_plot = 0,
  max_plot = 120
)

rambaud_age$plot
```

No discernable difference, so normal distribution is selected, as age
usually follows a normal distribution.

```{r fit-rambaud-ht, warning=FALSE, message=FALSE}
#| warning: false
rambaud_ht <- fit_cov_wrapper(
  cov_data = cov_data_rambaud,
  cov_name = "HT",
  min_plot = 140,
  max_plot = 200
)

rambaud_ht$plot
```

No discernable difference. Log-normal distribution is selected.

### Simulation of covariates

Covariates will be sampled from the following distributions :

-   CREAT : Log-normal distribution with a minimum 10 micromol/L with
    mean `r signif(exp(rambaud_creat$lnorm_par["meanlog"]),3)`
    micromol/L and standard deviation
    `r signif(rambaud_creat$lnorm_par["sdlog"],3)` micromol/L
-   WT : Normal distribution with resampling of negative values with
    mean `r signif(exp(rambaud_wt$lnorm_par["meanlog"]),3)` kg, and
    standard deviation
    `r signif((rambaud_wt$lnorm_par["sdlog"]), 3) * 100` %
-   AGE : Normal distribution with resampling of negative values with
    mean `r signif(exp(rambaud_age$norm_par["mean"]),3)` years, and
    coefficient of variation
    `r signif((rambaud_age$norm_par["sd"]), 3) * 100` %
-   HT : Normal distribution with resampling of negative values with
    mean `r signif(exp(rambaud_ht$lnorm_par["meanlog"]),3)` cm, and
    standard deviation
    `r signif((rambaud_ht$lnorm_par["sdlog"]), 3) * 100` %
    
The proportion of males is 0.794 in the original article, so the number of males is 496 and the number of females is 129 to add up to 625.

```{r sim-cov-rambaud, warning=FALSE, message=FALSE}
correlated_simulation <- function(n, means, cov_matrix) {
  set.seed(1991)
  collected <- matrix(NA, 0, length(means))
  colnames(collected) <- names(means)

  while (nrow(collected) < n) {
    batch <- MASS::mvrnorm(n = n, mu = means, Sigma = cov_matrix)
    colnames(batch) <- names(means)

    valid <- batch[, "AGE"] >= 18 &
             batch[, "CREAT_log"] < 5.991465 # log(400) corresponding to CRCL of 10 mL/min for a patient of 24 kg and 116 years

    batch_valid <- batch[valid, , drop = FALSE]
    collected <- rbind(collected, batch_valid)
  }

  collected[1:n, , drop = FALSE]
}

means <- c(
  CREAT_log = rambaud_creat$lnorm_par["meanlog"],
  WT_log      = rambaud_wt$lnorm_par["meanlog"], 
  AGE     = rambaud_age$norm_par["mean"],
  HT_log      = rambaud_ht$lnorm_par["meanlog"]
)

sds <- c(
  CREAT_log = rambaud_creat$lnorm_par["sdlog"],
  WT_log     = rambaud_wt$lnorm_par["sdlog"],
  AGE = rambaud_age$norm_par["sd"],
  HT_log     = rambaud_ht$lnorm_par["sdlog"]
)

covariates <- c("CREAT_log", "WT_log", "AGE", "HT_log")

names(means) <- covariates

# Correlation and Covariance Matrix
cor_rambaud_M <- cor_matrix_Rambaud_M[covariates, covariates]
cov_matrix_M <- diag(sds) %*% cor_rambaud_M %*% diag(sds)
eigen(cov_matrix_M)$values

cor_rambaud_F <- cor_matrix_Rambaud_F[covariates, covariates]
cov_matrix_F <- diag(sds) %*% cor_rambaud_F %*% diag(sds)
eigen(cov_matrix_F)$values

# Simulate
sim_data_M <- correlated_simulation(n = 496, means = means, cov_matrix = cov_matrix_M)
sim_data_M <- as_tibble(sim_data_M) %>%
  mutate(SEX = 0)

sim_data_F <- correlated_simulation(n = 129, means = means, cov_matrix = cov_matrix_F)
sim_data_F <- as_tibble(sim_data_F) %>%
  mutate(SEX = 1)

sim_data <- rbind(sim_data_M, sim_data_F)

# Put the covariates together
sim_cov_rambaud <- tibble(
  Paper = "Rambaud_2020",
  ID_within_paper = 1:n_patient,
  ICU = 1,
  BURN = 0,
  OBESE = 0,
  CREAT_log = sim_data$CREAT_log,
  WT_log   = sim_data$WT_log,
  AGE  = sim_data$AGE,
  HT_log  = sim_data$HT_log,
  SEX = sim_data$SEX)

sim_cov_rambaud <- sim_cov_rambaud %>%
  mutate(CREAT = exp(CREAT_log),
         HT = exp(HT_log),
         WT = exp(WT_log))

summary_sim_cov_rambaud <- sim_cov_rambaud |> 
  pivot_longer(cols = c(ICU,BURN,OBESE,CREAT,WT,AGE,HT),
               names_to = "Covariate") |> 
  summarise(.by = c(Covariate,Paper),
            Min = min(value),
            Q1 = quantile(value, 0.25),
            Median = quantile(value, 0.5),
            Q3 = quantile(value, 0.75),
            Max = max(value))
```

```{r rambaud-quantile-table, warning=FALSE, message=FALSE}
cov_rambaud_compare <- summary_sim_cov_rambaud |>
  rename_with(~ paste0(.x, "_sim"), -one_of("Covariate", "Paper")) |>
  left_join(rename_with(
    cov_data_rambaud,
    ~ paste0(.x, "_true"),
    -one_of("Covariate", "Paper", "Unit")
  )) |> 
  relocate(Covariate, .after = Paper) |> 
  relocate(Unit, .after = Covariate)

cov_rambaud_compare |>
  gt() |>
  fmt_scientific() |> 
    tab_spanner(columns = starts_with("Min"),
              label = "Min") |> 
      tab_spanner(columns = starts_with("Q1"),
              label = "Q1") |> 
      tab_spanner(columns = starts_with("Median"),
              label = "Median") |> 
      tab_spanner(columns = starts_with("Q3"),
              label = "Q3") |> 
  tab_spanner(columns = starts_with("Max"),
              label = "Max")
```

```{r plot-cov-rambaud, warning=FALSE, message=FALSE}
plot_sim_cov_wrapper(cov_data_rambaud,"CREAT",sim_cov_rambaud)
plot_sim_cov_wrapper(cov_data_rambaud,"WT",sim_cov_rambaud)
plot_sim_cov_wrapper(cov_data_rambaud,"AGE",sim_cov_rambaud)
plot_sim_cov_wrapper(cov_data_rambaud,"HT",sim_cov_rambaud)
```

CREAT is converted from micromol/L to mg/dL.

```{r reconvert_creat_rambaud, warning = FALSE, message = FALSE}
sim_cov_rambaud <- sim_cov_rambaud %>%
  mutate(
    BSA = (WT^0.425 * (HT^0.725) * 0.007184),
    CREAT = CREAT/88.4,
    BMI = WT / (HT/100)^2,
    OBESE = ifelse(BMI > 30, 1, 0)
  ) %>%
  dplyr::select(Paper, ID_within_paper, ICU, BURN, OBESE, CREAT, WT, BSA, BMI, AGE, SEX, HT)

# Calculate the simulated correlation matrices separately for the two sexes to compare with the original
sim_cov_rambaud_M <- sim_cov_rambaud %>% filter(SEX == 0)
sim_cov_rambaud_F <- sim_cov_rambaud %>% filter(SEX == 1)

cor_sim_M <- cor(sim_cov_rambaud_M %>% dplyr::select(WT, CREAT, AGE, HT, BMI), use = "complete.obs")
cor_sim_F <- cor(sim_cov_rambaud_F %>% dplyr::select(WT, CREAT, AGE, HT, BMI), use = "complete.obs")

ggcorrplot(cor_matrix_Rambaud_F, lab = TRUE, title = "Rambaud original (females)")
Rambaud_F <- ggcorrplot(cor_sim_F, lab = TRUE, title = "Rambaud simulated (females)")
Rambaud_F
ggcorrplot(cor_matrix_Rambaud_M, lab = TRUE, title = "Rambaud original (males)")
Rambaud_M <- ggcorrplot(cor_sim_M, lab = TRUE, title = "Rambaud simulated (males)")
Rambaud_M
```

## Final covariate table

Make the final covariate dataset with 2500 sets of covariates in a way
that in each 100 sets, we have 25-25 from each covariate's cohort, as
later, each dosing regimen will be simulated for 100 sets of covariates,
and it is important that the 4 cohorts are equally represented.

Also, for the concentration simulation part, we have to know before
applying the models, if a patient needs a different regiment due to his
renal failure or not. That is why an additional IR column is added where
1 indicates a renal failure patient who needs a dose adjustement. For
Carlier, the MDRD equation is used, and for Fournier, the Cockroft and
Gault. For Mellon and Rambaud, all patients have 0 in the IR column, as
it is not explicited in the articles that a dose adjustement was applied
to renal failure patients.

```{r final_table, warning=FALSE, message=FALSE}
# Patchwork simulated correlation plots
corr_simulated <- (Carlier_F + Fournier_F + Mellon_F + Rambaud_F +
  Carlier_M + Fournier_M + Mellon_M + Rambaud_M) + 
  plot_layout(ncol = 4, nrow = 2, byrow = TRUE)
ggsave(filename = here("Figures/S5.jpg"),
       plot = corr_simulated,
       width = 14, height = 8, dpi = 300)

# Row-bind the datasets to have 25 subjects from each in a sample of 100
datasets <- list(sim_cov_carlier, sim_cov_fournier, sim_cov_mellon, sim_cov_rambaud)

interleave_rows <- function(datasets, chunk_size) {
  n_chunks <- nrow(datasets[[1]]) / chunk_size
  interleaved <- do.call(rbind, lapply(1:n_chunks, function(i) {
    do.call(rbind, lapply(datasets, function(dataset) {
      dataset[((i - 1) * chunk_size + 1):(i * chunk_size), ]
    }))
  }))
  return(interleaved)
}

COV <- interleave_rows(datasets, chunk_size = 25)

# Convert Paper column to MODEL_COHORT column which indicates the model that is used to simulated the covariates
COV <- COV %>%
  mutate(MODEL_COHORT = case_when(
    Paper == "Carlier_2013" ~ "CARLIER",
    Paper == "Fournier_2018" ~ "FOURNIER",
    Paper == "Mellon_2020" ~ "MELLON",
    Paper == "Rambaud_2020" ~ "RAMBAUD")
  )

# Adding ID
COV$ID <- seq(1:2500)

# CRCL calculation functions
calculate_MDRD <- function(AGE, CREAT, SEX) {
  # Mutliply by 0.742 for females
    sex_factor <- ifelse(SEX == 0, 1, 0.742)
    eGFR <- 175 * (CREAT^(-1.154)) * AGE^(-0.203) * sex_factor
    return(eGFR)
}

calculate_CG <- function(AGE, CREAT, SEX, WT) {
  sex_factor <- ifelse(SEX == 0, 1, 0.85)
  eGFR <- ((140 - AGE) * WT * sex_factor) / (CREAT * 72)
  return(eGFR)
}

# Add 1 for dose adjustement for CRCL < 30 mL/min for Carlier and Fournier
COV <- COV %>%
  rowwise() %>%
  mutate(
    CRCL = case_when(
      MODEL_COHORT == "CARLIER" ~ calculate_MDRD(AGE, CREAT, SEX) * (BSA / 1.73),
      MODEL_COHORT == "FOURNIER" ~ calculate_CG(AGE, CREAT, SEX, WT),
      TRUE ~ NA_real_
    ),
    IR = case_when(
      MODEL_COHORT == "CARLIER" & CRCL < 30 ~ 1,
      MODEL_COHORT == "FOURNIER" & CRCL < 30 ~ 1,
      MODEL_COHORT %in% c("MELLON", "RAMBAUD") ~ 0,
      TRUE ~ 0
    )
  ) %>%
  ungroup()

COV <- COV %>%
  dplyr::select(ID, MODEL_COHORT, WT, CREAT, BURN, ICU, OBESE, BSA, AGE, SEX, HT, IR)

# Export to csv
write.csv(COV, here("Data/COV.csv"), quote = F, row.names = F)

# Check correlation (Pearson)
cor_sim <- cor(COV %>% dplyr::select(WT, CREAT, BURN, ICU, OBESE, AGE, SEX, HT), use = "complete.obs")
ggcorrplot(cor_sim, lab = TRUE)
```
